%File: ~/OOP/domain/load/pattern/LoadPattern.tex
%What: "@(#) LoadPattern.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$domain/load/pattern/LoadPattern.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class LoadPattern: public DomainComponent

\noindent {\bf Class Hierarchy}
\indent TaggedObject
\indent MovableObject
\indent\indent DomainComponent
\indent\indent\indent {\bf LoadPattern}

\noindent {\bf Description} 
\indent The LoadPattern class is a concrete base class. A
//! LoadPattern is a container class for Load and SFreedom\_Constraint
//! objects. Each LoadPattern object is assocaited with a TimeSeries
//! object which, for a given pseudo time, will return the appropriate
//! load factor to be applied to th load in the LoadPattern.

\noindent {\bf Class Interface}
\indent // Constructors 
\indent {\em LoadPattern(int tag);} 
\indent {\em LoadPattern(int tag, int classTag);}
\indent // Destructor 
\indent {\em virtual \f$\tilde{ }\f$LoadPattern();}
\indent // Public Methods 
\indent {\em virtual void setTimeSeries(TimeSeries *theSeries);}
\indent {\em virtual void setDomain(Domain *theDomain);}
\indent {\em  // Public Methods to add loads}
\indent {\em  virtual bool addNodalLoad(NodalLoad *);}
\indent {\em  virtual bool addElementalLoad(ElementalLoad *);}
\indent {\em  virtual bool addSFreedom\_Constraint(SFreedom\_Constraint *);} 
\indent {\em  virtual NodalLoadIter     \&getNodalLoads(void);}
\indent {\em  virtual ElementalLoadIter \&getElementalLoads(void);}
\indent {\em  virtual SFreedom\_ConstraintIter \&getSPs(void);}
\indent {\em  // Public Methods to remove loads}
\indent {\em  virtual NodalLoad *removeNodalLoad(int tag);}
\indent {\em  virtual ElementalLoad *removeElementalLoad(int tag);}
\indent {\em  virtual SFreedom\_Constraint *removeSFreedom\_Constraint(int tag);}
\indent {\em  // Public Methods to apply loads}
\indent {\em  virtual void applyLoad(double pseudoTime = 0.0);}
\indent {\em  virtual void setLoadConstant(void);}
\indent {\em  // Public Methods for o/p}
\indent {\em  virtual int sendSelf(int commitTag, Channel \&theChannel);}
\indent {\em  virtual int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);}
\indent {\em  virtual void Print(std::ostream \&s, int flag =0);}


\noindent {\bf Constructor} 
\indent {\em LoadPattern(int tag);} 
//! The integer \p tag is passed to the DomainComponent classes
//! constructor. Creates three ArrayOftaggedObjects objects to store
//! pointers to the NodalLoad, ElementalLoad and SFreedom\_Constraints and three
//! iters. If not enough memory is available for these objects an error
//! message is printed and the program is terminated.  

\indent {\em LoadPattern(int tag, classTag);} 
//! This is the constructor provided for subclasses to use. The integers
\p tag and \p classTag are passed to the DomainComponent classes
//! constructor. 

\noindent {\bf Destructor}
\indent {\em virtual \f$\tilde{ }\f$LoadPattern();} 
//! Invokes the destructor on the TimeSeries object. Also invokes the
//! destructor on any objects created in the constructor for storage of
//! the pointers and for iters. It does not invoke the destructor on these
//! objects, the Domain object is responsible for doing this.

\noindent {\bf Public Methods}
\indent {\em virtual void setTimeSeries(TimeSeries *theSeries);}
//! If a TimeSeries object is associated with the pattern, the destructor
//! is invoked on that TimeSeries object. It then sets the TimeSeries object
//! associated with the LoadPattern to \p theSeries.

\indent {\em virtual void setDomain(Domain *theDomain);} 
//! If any loads or constraint objects exist in the LoadPattern, the
//! LoadPattern will invoke setDomain() on those objects. Finally
//! invokes the DomainComponent classes setDomain() method.

\indent {\em  virtual bool addNodalLoad(NodalLoad *theLoad);}
//! Adds the NodalLoad object pointed to by \p theLoad to the
//! LoadPattern. If the LoadPattern could add the pointer to its storage
//! object for nodal loads, it will invoke setDomain()
//! and setLoadPattern() on the load object if a Domain has been set.

\indent {\em  virtual bool addElementalLoad(ElementalLoad *theLoad);}
//! Adds the ElementalLoad pointed to by \p theLoad to the
//! LoadPattern. If the LoadPattern could add the pointer to its storage
//! object for elemental loads, it will invoke setDomain()
//! and setLoadPattern() on the load object if a Domain has been set.

\indent {\em  virtual bool addSFreedom\_Constraint(SFreedom\_Constraint *theSp);} 
//! Adds the SFreedom\_Constraint pointed to by \p theSp to the
//! LoadPattern. If the LoadPattern could add the pointer to its storage
//! object for single-point constraints, it will invoke setDomain()
//! and setLoadPattern() on the constraint object if a Domain has
//! been set.

\indent {\em  virtual NodalLoadIter     \&getNodalLoads(void);}
//! Returns an iter to the nodal loads in the LoadPattern.

\indent {\em  virtual ElementalLoadIter \&getElementalLoads(void);}
//! Returns an iter to the elemental loads in the LoadPattern.

\indent {\em  virtual SFreedom\_ConstraintIter \&getSPs(void);} 
//! Returns an iter to the single-point constraints in the LoadPattern.
    
\indent {\em  virtual NodalLoad *removeNodalLoad(int tag);}
//! To remove the nodal load whose identifier is given by \p tag from
//! the LoadPattern and sets the laods associated Domain object to
\f$0\f$. Returns a pointer to the load if succesfully removed, otherwise
\f$0\f$ is returned. 

\indent {\em  virtual ElementalLoad *removeElementalLoad(int tag);}
//! To remove the elemental load whose identifier is given by \p tag from
//! the LoadPattern and set the loads associated Domain object to
\f$0\f$. Returns a pointer to the load if succesfully removed, otherwise
\f$0\f$ is returned. 

\indent {\em  virtual SFreedom\_Constraint *removeSFreedom\_Constraint(int tag);} 
//! To remove the single-point constraint whose identifier is given by \p tag from
//! the LoadPattern and st its associated Doman object to \f$0\f$. Returns a
//! pointer to the load if succesfully removed, otherwise \f$0\f$ is
//! returned. 

\indent {\em  virtual void applyLoad(double pseudoTime = 0.0);}
//! To apply the load for the pseudo time \p pseudoTime. From the
//! associated TimeSeries object the LoadPattern will obtain a current
//! load factor for the pseudo time. It will then invoke {\em
//! applyLoad(load factor)} on the loads and {\em applyConstraint(load
//! factor)} on the single-point constraints in the LoadPattern. If {\em
//! setLoadConstant()} has been invoked, the saved load factor is used and
//! no call is made to the TimeSeries object. If no TimeSeries is
//! associated with the object a load factor of \f$0.0\f$ is used.

\indent {\em  virtual void setLoadConstant(void);} 
//! Marks the LoadPattern as being constant. Subsequent calls to {\em
//! applyLoad()} will use the current load factor.

\indent {\em  virtual int sendSelf(int commitTag, Channel
\&theChannel);}
//! Creates a vector of size 4 into which it places the current load
//! factor, the mark indicating whether LoadPattern is constant, and the
//! database tag and class tag of the TimeSeris object. Invokes {\em
//! sendVector()} on the Channel object and sendSelf() on the
//! TimeSeries object.

\indent {\em  virtual int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);}
//! Does the mirror image of sendSelf().

\indent {\em  virtual void Print(std::ostream \&s, int flag =0);}
//! Invokes {\em Print(s, flag)} on the TimeSeries, NodalLoads,
//! ElementalLoads and SFreedom\_Constraints. 
