%File: ~/OOP/element/node/NodalLoad.tex
%What: "@(#) NodalLoad.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/domain/node/NodalLoad.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class NodalLoad: public Load

\noindent {\bf Class Hierarchy}
\indent TaggedObject
\indent MovableObject
\indent\indent DomainComponent
\indent\indent\indent Load
\indent\indent\indent\indent {\bf NodalLoad}

\noindent {\bf Description}
\indent NodalLoads are loads acting on Nodes. The public methods are
//! all declared as virtual to allow subclasses to be introduced for the
//! provision of time varying loads. Each NodalLoad object is associated
//! with a single Node object and has a Vector object corresponding to the
//! load acting on this Node object as a result of the NodalLoad.

\noindent {\bf Class Interface}
\indent\indent // Constructors
\indent\indent {\em NodalLoad(int tag, int node, const Vector \&load, bool
//! isLoadConstant = false);} 
\indent\indent {\em NodalLoad(int tag, int node, int classTag);}
\indent\indent {\em NodalLoad(int classTag);}
\indent\indent // Destructor
\indent\indent {\em virtual \f$\tilde{ }\f$ NodalLoad();}
\indent\indent // Public Methods
\indent\indent {\em virtual void setDomain(Domain *newDomain);}
\indent\indent {\em virtual int getNodeTag(void) const;}
\indent\indent {\em virtual void applyLoad(double loadFactor);}
\indent\indent // Public Methods for Output
\indent\indent {\em virtual int sendSelf(int commitTag, Channel \&theChannel); }
\indent\indent {\em virtual int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker); }
\indent\indent {\em virtual void Print(std::stream \&s, int flag = 0);}


\noindent {\bf Constructors}
\indent {\em NodalLoad(int tag, int node, const Vector \&load, bool
//! isLoadConstant = false);}
//! To create a NodalLoad object with tag \p tag acting on Node \p node with a
//! reference load given by \p load. \p tag and {\em LOAD\_TAG\_NodalLoad} (defined in
\f$<\f$classTags.h\f$>\f$)are passed to the Load constructor.  A new vector object is created using
//! the vector \p load as the argument. (THIS MAY CHANGE - may associate
//! the load Vector with \p load, allowing identical loads on diff
//! nodes to be created without the need for each to have its own
//! vector) If no enough memory is available an error message is printed
//! and the program terminates. The boolean \p isLoadConstant is used
//! to indicate whether the value of the load applies to the Node is
//! independent of the load factor. 

\indent {\em NodalLoad(int tag, int node, int classTag);}
//! Provided for subclasses, which may which to provide the abstraction of
//! time varying nodal loads. The integers \p tag and \p classTag
//! are passed to the Load constructor.

\indent {\em NodalLoad(int classTag);}
//! Provided so that a FEM\_ObjectBroker can construct an object. \f$0\f$ and
\p classTag are passed to the Load classes constructor. The data
//! for the object is filled in when recvSelf() is invoked on the
//! object.

\noindent {\bf Destructor}
\indent {\em virtual \f$\tilde{ }\f$ NodalLoad();} 
//! If a Vector for the load was constructed, the destructor invokes
//! delete on this Vector object.


\noindent {\bf Public Methods}
\indent {\em virtual void setDomain(Domain *newDomain);}
//! To set the associated Domain object. First checks to ensure that a
//! Node object with the tag exists in the Domain. It sets the pointer for
//! it's associated Node object to point to this object, and then 
//! invokes the DomainComponent classes setDomain() method. If the
//! Node does not exist in the Domain a warning message and {\em
//! setDomain()} is not invoked.

\indent {\em virtual int getNodeTag(void) const;}
//! Returns the integer \p node passed in the constructor.

\indent {\em virtual void applyLoad(double timeStep = 0.0, double
//! loadFactor = 1.0);}
//! To it's associated Node it invokes addUnbalancedLoad() with it's
//! copy of the Vector object \p load and a factor of \p loadFactor
//! if \p isLoadConstant was specified as \p false in the constructor or
\f$1\f$ if it was specified as \p true. Warning
//! messages are printed, if no Domain has been associated with the
//! NodalLoad object or no Node with a tag \p node exists in the
//! Domain. For efficiency reasons, the NodalLoad object keeps a pointer
//! to it's associated Node object. The time \p timeStep has no
//! influence on the load added.

\indent {\em virtual int sendSelf(int commitTag, Channel \&theChannel); }
//! Determines its database tag. The object then sends it's tag, {\em
//! node} and size of load Vector to the Channel in an ID object. Then, if
\p load is not NULL, it sends it's copy of the \p load
//! Vector. Returns \f$0\f$ if successful, a negative number if the Channel
//! failed to send the data.  

\indent {\em virtual int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker); }
//! The object first determines its database tag. It then invokes receives
//! an ID object from the Channel containing it's tag, \p node and size
//! of load Vector. If size of \p load is not \f$0\f$ it then receives it's
//! copy of the \p load Vector. Returns \f$0\f$ if successful, a negative 
//! number if the Channel failed to receive the data. 

\indent {\em virtual void Print(std::stream \&s, int flag = 0);}
//! Prints it's \p node and then prints the load Vector.

