%File: ~/OOP/graph/graph/Metis.tex
%What: "@(#) Metis.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/graph/partitioner/Metis.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class GraphPartitioner:

\noindent {\bf Class Hierarchy}
\indent GraphPartitioner
\indent\indent {\bf Metis}

\noindent {\bf Description}
\indent Metis is a GraphPartitioner. The Metis graph partitioner calls
//! procedures defined in the METIS library to partition the graph. METIS
//! is currently being developed by G.~Karypis and V.~Kumar at the
//! University of Minnesota. At the present time the Graph to be
//! partitioned MUST have the vertices labeled \f$0\f$ through \f$numVertex-1\f$.

//! The METIS library uses two integer arrays to represent the graph, {\em
//! xadj} and \p adjncy. \f$xadj(i)\f$ stores the location in \p adjncy
//! of the start of the \f$i\f$'th Vertices adjacent Vertices. \p adjncy
//! contains the tags of all the adjacent vertices. For example, the graph
//! which is represented by the following matrix \f$A\f$:


\f$\f$ A =
\left[
\begin{array}{ccccc}
1 & 0 & 1 & 1 & 0
1 & 1 & 0 & 0 & 0
0 & 1 & 1 & 0 & 0
0 & 0 & 0 & 1 & 1
1 & 1 & 0 & 0 & 1
\end{array}
\right] 
\f$\f$

//! is represented by:

\f$\f$
//! xadj =
\left[
\begin{array}{cccccccccccccc}
0 & 2 & 3 & 4 & 5 & 7
\end{array}
\right] 
\f$\f$

//! and

\f$\f$
//! adjncy =
\left[
\begin{array}{cccccccccccccc}
2 & 3 & 0 & 1 & 4 & 0 & 1
\end{array}
\right] 
\f$\f$

\noindent note that there is no space allocated for the diagonal
//! components.

\noindent {\bf Class Interface}
\indent\indent // Constructors
\indent\indent {\em Metis();}
\indent\indent {\em Metis(int pType, int mType, int coarsenTo, int 
//! rType, int ipType);}
\indent\indent // Destructor
\indent\indent {\em virtual~\f$\tilde{}\f$Metis();}
\indent\indent // Public Methods
\indent\indent {\em virtual int partition(Graph \&theGraph, int numPart) =0;}
\indent\indent {\em bool setDefaultOptions(void);}
\indent\indent {\em bool setOptions(int pType, int mType, int coarsenTo, int 
//! rTypem, int ipType);}
\indent\indent // Private Method
\indent\indent {\em bool checkOptions(void) const;}


\noindent {\bf Constructors}
\indent {\em Metis();}
//! To construct a Metis object which will use the default settings when
//! partitioning. 

\indent {\em Metis(int pType, int mType, int coarsenTo, int 
//! rTypem, int ipType);}
//! To construct a Metis object which will use the setting passed into the
//! constructor as options to metis's PMETIS() routine. {\em
//! checkOptions()} is invoked to ensure the settings are valid.

\noindent {\bf Destructor}
\indent {\em virtual~\f$\tilde{}\f$Metis();}

\noindent {\bf Public Methods }
\indent {\em virtual int partition(Graph \&theGraph, int numPart) =0;}
//! This is the method invoked to partition the graph into \p numPart
//! partitions. On completion of the routine each vertex will be assigned
//! a color \f$1\f$ through \p numPart, the color assigned indicating the
//! partition to which the vertex belongs. 

//! To partition a number of integer arrays are created, {\em options[5]},
{\em partition[numVertex+1]}, {\em xadj[numVertex+1]} and {\em
//! adjncy[2*numEdge]} (CURRENTLY ASSUMING GRAPH IS SYMMETRIC - THIS MAY
//! CHANGE \& xadj and partition 1 LARGER THAN REQUIRED). If not enough
//! memory is available for the arrays, a warning message is printed and
\f$-2\f$ is returned. The data for \p xadj and \p adjncy are
//! determined from the Vertices of the Graph by iterating over each
//! Vertex from \f$0\f$ through \p numVertex \f$-1\f$. If default options are
//! specified {\em options[0]} is set to \f$0\f$, otherwise \f$1\f$ with {\em
//! options[1:4] = coarsenTo, mType, ipType, rType}. if \p pType equals
\f$1\f$ \p PMETIS is called, otherwise \p KMETIS is called. Both are
//! called with the following arguments: {\em numVertex, xadj,adjncy, 0,
0, \&weightFlag, options, numPart, \&numbering, \&edgecut, partition} 
//! The colors of the partitions are then set equal to the color indicated
//! in \p partition.  The integer arrays are destroyed and \f$0\f$
//! returned.

\indent {\em bool setDefaultOptions(void);}
//! Sets the default options.

\indent {\em bool setOptions(int pType, int mType, int coarsenTo, int 
//! rType, int ipType);} 
//! Sets the options for the partitioning to those passed as
//! arguments. Then invokes checkOptions() to see if the options are
//! valid. HOW ABOUT REFERRINGR TO MANUAL TO SEE WHAT OPTIONS MEAN.

\noindent {\bf Private Method}
\indent {\em bool checkOptions(void) const;}
//! If options are not valid sets the default options. EXPAND ON VALID
//! OPTIONS OR REFER TO METIS MANUAL.
