%File: ~/OOP/analysis/integrator/TransientIntegrator.tex
%What: "@(#) TransientIntegrator.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/analysis/integrator/TransientIntegrator.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class TransientIntegrator: public Integrator

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent Integrator
\indent\indent\indent IncrementalIntegrator
\indent\indent\indent\indent {\bf TransientIntegrator}
\indent\indent\indent\indent\indent Newmark
\indent\indent\indent\indent\indent HHT
\indent\indent\indent\indent\indent Wilson-\f$\Theta\f$

\noindent {\bf Description} 
\indent TransientIntegrator is an abstract subclass of IncrementalIntegrator.
//! A subclass of it is used when performing a nonlinear transient
//! analysis of the problem using a direct integration method. The
//! TransientIntegrator class redefines the formTangent() method of
//! the IncrementalIntegrator class and it defines a new method {\em
//! newStep()} which is invoked by the DirectIntegrationAnalysis class at
//! each new time step.

//! In nonlinear transient finite element problems we seek a solution
(\f$\U\f$, \f$\dot \U\f$, \f$\ddot \U\f$) to the nonlinear vector function

\begin{equation}
\R(\U,\Ud, \Udd) = \P(t) - \F_I(\Udd) - \F_R(\U, \Ud) = \zero
\label{femGenForm}
\end{equation}


//! The most widely used technique for solving the transient non-linear 
//! finite element equation, equation~\ref{femGenForm}, is to use an
//! incremental direct integration scheme. In the incremental formulation,
//! a solution to the equation is sought at successive time steps \f$\Delta
//! t\f$ apart.  

\begin{equation}
\R(\U_{n \Delta t},\Ud_{n \Delta t}, \Udd_{n \Delta t}) = \P(n \Delta t) -
\F_I(\Udd_{n \Delta t}) - \F_R(\U_{n \Delta t}, \Ud_{n \Delta t})
\label{fullTimeForm}
\end{equation}

//! For each time step, t, the integration schemes provide two operators,
\f$\I_1\f$ and \f$\I_2\f$, to relate the velocity and accelerations at the 
//! time step as a function of the displacement at the time step and the
//! response at previous time steps: 

\begin{equation} 
\dot \U_{t} = {\I}_1 (\U_t, \U_{t-\Delta t}, \dot \U_{t-\Delta t},
\ddot \U_{t - \Delta t}, \U_{t - 2\Delta t}, \dot \U_{t - 2 \Delta t}. ..., )
\label{I1}
\end{equation} 

\begin{equation} 
\ddot \U_{t} = {\I}_2 (\U_t, \U_{t-\Delta t}, \dot \U_{t-\Delta t},
\ddot \U_{t - \Delta t}, \U_{t - 2\Delta t}, \dot \U_{t - 2 \Delta t}. ..., )
\label{I2}
\end{equation} 

//! These allow us to rewrite equation~\ref{fullTimeForm}, in terms of a
//! single response quantity, typically the displacement:

\begin{equation}
\R(\U_t) = \P(t) - \F_I(\Udd_t) - \F_R(\U_t, \Ud_t)
\label{genForm}
\end{equation}

//! The solution of this equation is typically obtained using an iterative
//! procedure, i.e. making an initial prediction for \f$\U_{t}\f$,
//! denoted \f$\U_{t}^{(0)}\f$ a sequence of approximations \f$\U_{t}^{(i)}\f$,
\f$i=1,2, ..\f$ is obtained which converges (we hope) to the solution \f$\U_{t}\f$. The
//! most frequently used iterative schemes, such as Newton-Raphson,
//! modified Newton, and quasi Newton schemes, are based on a Taylor
//! expansion of equation~\ref{genForm} about \f$\U_{t}\f$:    

\begin{equation} 
\R(\U_{t}) = 
\R(\U_{t}^{(i)}) +
\left[ {\frac{\partial \R}{\partial \U_t} \vert}_{\U_{t}^{(i)}}\right]
\left( \U_{t} - \U_{t}^{(i)} \right) 
\end{equation}

\f$\f$
\R(\U_{t}) = 
 \P (t) 
 - \f_{I} \left( \ddot \U_{t}^{(i)} \right) -
\f_{R} \left( \dot \U_{t}^{(i)}, \U_{t}^{(i)} \right)
\f$\f$
\begin{equation} 
- \left[
   \M^{(i)} {\I}_2'
+  \C^{(i)} {\I}_1'
+ \K^{(i)}  \right]
 \left( \U_{t} - \U_{t}^{(i)} \right)
\label{femGenFormTaylor}
\end{equation} 

//! To start the iteration scheme, trial values for \f$\U_{t}\f$, \f$\dot
\U_{t} \f$ and \f$\ddot \U_{t} \f$ are required. These are
//! obtained by assuming \f$\U_{t}^{(0)} = \U_{t-\Delta t}\f$. The
\f$\dot \U_{t}^{(0)} \f$ and \f$\ddot \U_{t}^{(0)}\f$
//! can then be obtained from the operators for the integration scheme.

//! Subclasses of TransientIntegrators provide
//! methods informing the FE\_Element and DOF\_Group objects how to build
//! the tangent and residual matrices and vectors. They also provide the
//! method for updating the response quantities at the DOFs with
//! appropriate values; these values being some function of the solution
//! to the linear system of equations. 


\noindent {\bf Class Interface }
\indent // Constructor
\indent {\em TransientIntegrator(int classTag);}
\indent // Destructor 
\indent {\em virtual~ \f$\tilde{}\f$TransientIntegrator();}
\indent // Public Methods
\indent {\em virtual int formTangent(void);}
\indent {\em virtual int formEleResidual(FE\_Element *theEle);}
\indent {\em virtual int formNodUnbalance(DOF\_Group *theDof);}
\indent {\em virtual void newStep(double deltaT) =0;}

\noindent {\bf Constructor}
\indent {\em TransientIntegrator(int classTag);} 
//! The integer \p classTag is passed to the IncrementalIntegrator
//! classes constructor.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$TransientIntegrator();} 
//! Does nothing.

\noindent {\bf Public Methods}
\indent {\em virtual int formTangent(void);}
//! Invoked to form the structure tangent matrix. The method is rewritten
//! for this class to include inertia effects from the nodes. The method
//! iterates over both the FE\_Elements and DOF\_Groups invoking methods
//! to form their contributions to the \f$A\f$ matrix of the LinearSOE and
//! then adding these contributions to the \f$A\f$ matrix. The method performs
//! the following:
\begin{tabbing}
//! while \= \+ while \= while \= \kill
//! theSysOfEqn.zeroA();
//! DOF\_EleIter \&theDofs = theAnalysisModel.getDOFs();
//! while((dofGroupPtr = theDofs()) \f$\neq\f$ 0) \+
//! dofGroupPtr-\f$>\f$formTangent(theIntegrator);
//! theSOE.addA(dofGroupPtr-\f$>\f$getTangent(this),
//! dofGroupPtr-\f$>\f$getID()) \-
//! FE\_EleIter \&theEles = theAnalysisModel.getFEs();
//! while((elePtr = theEles()) \f$\neq\f$ 0) \+
//! theSOE.addA(elePtr-\f$>\f$getTangent(this),
//! elePtr-\f$>\f$getID(), \f$1.0\f$)
\end{tabbing}
\noindent Returns \f$0\f$ if successful, otherwise a \f$-1\f$ if an error occurred while
//! trying to add the stiffness. The two loops are introduced for the
//! FE\_Elements, to allow for efficient parallel programming when the
//! FE\_Elements are associated with a ShadowSubdomain. 

{\em virtual int formEleResidual(FE\_Element *theEle);}
//! Called upon by the FE\_Element \p theEle to determine it's
//! contribution to the rhs of the equation. The following are invoked
//! before \f$0\f$ is returned.
\begin{tabbing}
//! while \= \+ while \= while \= \kill
//! theEle-\f$>\f$zeroResidual()
//! theEle-\f$>\f$addRIncInertiaToResid()
\end{tabbing}

{\em virtual int formNodUnbalance(DOF\_Group *theDof);}
//! Called upon by the DOF\_Group \p theDof to determine it's
//! contribution to the rhs of the equation. The following are invoked
//! before \f$0\f$ is returned.
\begin{tabbing}
//! while \= \+ while \= while \= \kill
//! theDof-\f$>\f$zeroUnbalance()
//! theDof-\f$>\f$addPIncInertiaToUnbalance()
\end{tabbing}

{\em virtual int newStep(double deltaT) =0;}
//! Invoked to inform the integrator that the transient analysis is
//! proceeding to the next time step. To return \f$0\f$ if successful, a
//! negative number if not.




