%File: ~/OOP/graph/numberer/RCM.tex
%What: "@(#) RCM.tex, revA"


\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/graph/numberer/RCM.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class RCM: public GraphNumberer;

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent GraphNumberer
\indent\indent\indent {\bf RCM}

\noindent {\bf Description}
\indent RCM is a subclass of GraphNumberer which performs the
//! numbering using the reverse Cuthill-McKee numbering algorithm.

\noindent {\bf Class Interface }
\indent\indent // Constructor
\indent\indent {\em RCM(bool GPS = true);}
\indent\indent // Destructor
\indent\indent {\em ~\f$\tilde{}\f$RCM();}
\indent\indent // Public Methods
\indent\indent {\em const ID \&number(Graph \&theGraph, int
//! lastVertexTag = -1) =0;}
\indent\indent {\em const ID \&number(Graph \&theGraph, const ID
\&startVertices) =0;}
\indent\indent {\em int sendSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);}
\indent\indent {\em int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker); }

\noindent {\bf Constructor}
\indent {\em RCM(bool GPS = true);}
//! The integer \p classTag is passed to the MovableObject classes
//! constructor. The flag \p GPS is used to mark whether the
//! Gibbs-Poole-Stodlmyer algorithm is used to determine a starting vertex
//! when no starting vertex is given.

\noindent {\bf Destructor}
\indent {\em virtual~\f$\tilde{}\f$RCM();}
//! Invokes the destructor on any ID object created when number() is
//! invoked.

\noindent {\bf Public Methods}
\indent {\em const ID \&number(Graph \&theGraph, int
//! lastVertex = -1) =0;}
//! If the present ID used for the result is not of size equal to the
//! number of Vertices in \p theGraph, it deletes the old and
//! constructs a new ID. Starts by iterating through the Vertices of the
//! graph setting the \p tmp variable of each to \f$-1\f$. The Vertices are
//! then numbered using a depth first sort of the Graph, with each
//! unmarked Vertex in the Graph at a distance \f$d\f$ from starting Vertex
//! being placed in the d'th level set. As this is RCM, the Vertices in
//! level set \f$n\f$ are assigned a higher number than those in level set
\f$n+1\f$ with the \p tmp variable of the starting Vertex being
//! assigned \p numVertices \f$-1\f$. The \p tags of the Vertices are
//! placed into the ID at location given by their \p tmp variable. These
//! are replaced with the \p ref variable of each Vertex, which is
//! returned on successful completion. 


//! The Vertex chosen as the starting Vertex is the one whose tag is given
//! by \p lastVertex. If this is \f$-1\f$ or the Vertex corresponding to
\p lastVertex does not exist then another Vertex is chosen. If the
\p GPS flag in constructor is \p false the first Vertex from the
//! Graphs VertexIter is used; if \p true a RCM numbering using the
//! first Vertex from the VertexIter is performed and the Vertices in the
//! last level set are then used to create an ID \p lastVertices with
//! which {\em number(theGraph, lastVertices)} can be invoked to determine
//! the numbering.


\indent {\em const ID \&number(Graph \&theGraph, const ID
\&startVertices) =0;}
//! This method is invoked to determine the best starting Vertex for a RCM
//! using a Vertex whose tag is in \p lastVertices. To do a RCM
//! numbering is performed using each of the Vertices in {\em
//! startVertices} as the Vertex in level set \f$0\f$. The Vertex which
//! results in the numbering with the smallest profile is chosen as 
//! the starting Vertex. The RCM algorithm outlined above is then called
//! with this starting Vertex.

{\em int sendSelf(Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);}
//! Returns \f$0\f$.

{\em int recvSelf(Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker); }
//! Returns \f$0\f$.

