%File: ~/recorder/MaxNodeDispRecorder.tex
%What: "@(#) MaxNodeDispRecorder.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/recorder/MaxNodeDispRecorder.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class MaxNodeDispRecorder: public Recorder

\noindent {\bf Class Hierarchy}
\indent Recorder
\indent\indent {\bf MaxNodeDispRecorder}

\noindent {\bf Description}
\indent The MaxNodeDispRecorder class is used to store information
//! about the absolute maximum nodal displacement at a number of specified
//! Nodes for a specified degree of freedom during an analysis.

\noindent {\bf Class Interface}
\indent // Constructor
\indent {\em MaxNodeDispRecorder(int dof, const ID \&nodeTags, Domain
\&theDomain);} 
\indent // Destructor
\indent {\em \f$\tilde{ }\f$MaxNodeDispRecorder();}
\indent // Public Methods
\indent {\em int record(int commitTag);}
\indent {\em int playback(int commitTag);}
\indent {\em void restart(void);} 

\noindent {\bf Constructor}
\indent {\em MaxNodeDispRecorder(int dof, const ID \&nodeTags,
//! Domain \&theDomain);} 
//! Creates a \f$0\f$ Vector object of size equal to the size of {\em
//! nodeTags} to store the maximum nodal displacements
//! and creates an ID object equal to \p nodeTags to store the Node
//! tags.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$MaxNodeDispRecorder();} 
//! Does nothing.

\noindent {\bf Public Methods }
\indent {\em int record(int commitTag);}
//! For each node in \p nodeTags the value of the \p dof'th committed
//! displacement is obtained. If the absolute value of this is greater
//! than the value currently stored in the Vector of max displacements,
//! the value in the Vector is updated. If no Node exists in {\em
//! theDomain} with the tag or the Node does not have a \p dof'th
//! degree-of-freedom associated with it \f$0\f$ is entered in the
//! Vector. Returns \f$0\f$.

\indent {\em int playback(int commitTag);}
//! Prints to \p opserr the Vector containing the maximum absolute nodal
//! displacements. Note, at the end of the analysis, what is printed is
//! independent of \p commitTag. Returns \f$0\f$.

\indent {\em void restart(void);} 
//! Zeros the Vector of maximum nodal displacements.
