%File: ~/recorder/Recorder.tex
%What: "@(#) Recorder.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/recorder/Recorder.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class Recorder

\noindent {\bf Class Hierarchy}
\indent {\bf Recorder}

\noindent {\bf Description}
\indent The Recorder class is an abstract class which is introduced to allow
//! information to be saved during the analysis. The interface defines two
//! pure virtual methods record()} and {\em playback(). {\em
//! record()} is a method which is called by the Domain object during a
//! commit()}. The {\em playback() method can be called by the analyst after
//! the analysis has been performed.

\noindent {\bf Class Interface}
\indent // Constructor
\indent {\em Recorder();}
\indent // Destructor
\indent {\em virtual \f$\tilde{ }\f$Recorder();}
\indent // Public Methods
\indent {\em virtual int record(int commitTag) =0;}
\indent {\em virtual int playback(int commitTag) =0;}
\indent {\em virtual void restart(void) =0;} 

\noindent {\bf Constructor}
\indent {\em Recorder();} 
//! Does nothing.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$Recorder();} 
//! Does nothing.

\noindent {\bf Public Methods }
\indent {\em virtual int record(int commitTag) =0;}
//! Invoked by the Domain object after commit() has been invoked on all the
//! domain component objects. What the Recorder records depends on the
//! concrete subtype.

\indent {\em virtual int playback(int commitTag) =0;}
//! Invoked by the analyst after the analysis has been performed. What the
//! Recorder does depends on the concrete subtype.

\indent {\em virtual void restart(void) =0;} 
//! Invoked by the Domain object when revertToStart() is invoked on
//! the Domain object. What the Recorder does depends on the concrete subtype. 
