%File: ~/OOP/analysis/integrator/LoadPath.tex
%What: "@(#) LoadPath.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/analysis/integrator/LoadPath.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class LoadPath: public StaticIntegrator

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent Integrator
\indent\indent\indent IncrementalIntegrator
\indent\indent\indent\indent StaticIntegrator
\indent\indent\indent\indent\indent {\bf LoadPath}

\noindent {\bf Description} 
\indent LoadPath is a subclass of StaticIntegrator, it is
//! used to when performing a static analysis on the FE\_Model using a
//! user specified load path. The load path is specified in a Vector, {\em
//! path}, to the objects constructor and at each step in the analysis:

\[ 
\lambda_n^{(i)} - \lambda_{n-1} = path(n) - path(n-1)
\]

//! Knowing \f$\lambda_n^{(i)} = path(n)\f$ prior to each iteration, the \f$N+1\f$
//! unknowns in equation~\ref{staticFormTaylor}, is reduced to \f$N\f$ unknowns and
//! results in the following equation:

\begin{equation} 
\R(\U_{n}) = \lambda_n^{(i)} \P 
 - \f_{R}\left(\U_{n}^{(i)} \right) - 
\K_n^{(i)} 
(\U_{n} - \U_{n}^{(i)})  
\label{staticFormLoadPath}
\end{equation} 

\noindent {\bf Class Interface}
\indent // Constructors
\indent {\em LoadPath(Vector \&loadPath);} 
\indent {\em LoadPath();}
\indent // Destructor
\indent {\em \f$\tilde{ }\f$LoadPath();}
\indent // Public Methods
\indent {\em int newStep(void);}
\indent {\em int update(const Vector \&\f$\Delta U\f$);}
\indent // Public Methods for Output
\indent {\em int sendSelf(int commitTag, Channel \&theChannel);} 
\indent {\em int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);} 
\indent {\em int Print(OPS\_Stream \&s, int flag = 0);}

\noindent {\bf Constructors}
\indent {\em LoadPath(Vector \&loadPath);} 
//! The integer INTEGRATOR\_TAGS\_LoadPath (defined in
\f$<\f$classTags.h\f$>\f$) is passed to the StaticIntegrator classes
//! constructor. A vector object \p path is created which is a copy of
\p loadPath and an index into this vector is set to \f$0\f$.

\indent {\em LoadPath();} 
//! The integer INTEGRATOR\_TAGS\_LoadPath (defined in \f$<\f$classTags.h\f$>\f$) is
//! passed to the StaticIntegrator classes constructor. No vector object
//! is created. Provided for the FEM\_ObjectBroker to create a blank
//! object, recvSelf() should be invoked on this object.

\noindent {\bf Destructor}
\indent {\em \f$\tilde{ }\f$LoadPath();} 
//! Invokes the destructor on the vector \p path.

\noindent {\bf Public Methods}
\indent {\em int newStep(void);}
//! The object obtains the current value of \f$\lambda\f$ from the \p path
//! vector using the current index. The index is then incremented by
\f$1\f$. If the index is greater than the size of \p path, \f$\lambda\f$ is
//! set to \f$0\f$ and an error message is printed. It will then invoke
{\em applyLoadDomain(0.0, \f$\lambda\f$)} on the AnalysisModel
//! object. Returns \f$0\f$ if successful. A warning message is printed and a
\f$-1\f$ is returned if no AnalysisModel is associated with the object.

{\em int update(const Vector \&\f$\Delta U\f$);}
//! Invoked this causes the object to first increment the DOF\_Group
//! displacements with \f$\Delta U\f$, by invoking {\em incrDisp(\f$\Delta U)\f$}
//! on the AnalysisModel, and then to update the domain, by invoking {\em
//! updateDomain()} on the AnalysisModel. Returns \f$0\f$ if successful, a
//! warning message and a \f$-1\f$ is returned if no AnalysisModel is
//! associated with the object.


{\em int sendSelf(int commitTag, Channel \&theChannel); } 
//! Places the size if \p path and the index in an ID of size 2 and send this
//! ID. Then sends the Vector \p path. Returns \f$0\f$ if successful, a
//! warning message is printed and a \f$-1\f$ is returned if \p theChannel
//! fails to send the ID or the Vector. 

{\em int recvSelf(int commitTag, Channel \&theChannel, 
//! FEM\_ObjectBroker \&theBroker); } 
//! Receives in a ID of size 2 the size of the vector and current index.
//! Creates a new Vector and receives the Vector from the Channel. 
//! Returns \f$0\f$ if successful, a warning message is printed and a \f$-1\f$ is
//! returned if \p theChannel fails to receive the Vector or the ID.

{\em int Print(OPS\_Stream \&s, int flag = 0);}
//! The object sends to \f$s\f$ its type, the current value of \f$\lambda\f$.
