%File: ~/OOP/analysis/numberer/DOF_Numberer.tex
%What: "@(#) DOF_Numberer.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/analysis/numberer/DOF\_Numberer.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class DOF\_Numberer: public MovableObject

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent {\bf DOF\_Numberer}

\noindent {\bf Description} 
\indent The DOF\_Numberer class is a base class. Its purpose is
//! to define the interface common among all subclasses.  A DOF\_Numberer
//! object is responsible for assigning the equation numbers to the
//! individual dofs in each of the  DOF\_Groups in the AnalysisModel. The
//! base DOF\_Numberer uses a GraphNumberer object to first number the
//! DOF\_Groups, based on the ordering of the DOF\_Group objects, it
//! assigns the equation numbers to the individual
//! degrees-of-freedom. Subtypes may wish to implement the numbering in
//! a more efficient manner by using the FE\_Element and DOF\_Group objects
//! directly.


\noindent {\bf Class Interface}
\indent // Constructors
\indent {\em DOF\_Numberer(GraphNumberer \&theGraphNumberer);} 
\indent {\em DOF\_Numberer(int classTag);} 
\indent {\em DOF\_Numberer();}
\indent // Destructor
\indent {\em virtual~ \f$\tilde{}\f$DOF\_Numberer();}
\indent // Public Methods
\indent {\em void setLinks(AnalysisModel \&theModel);} 
\indent {\em virtual int numberDOF(int lastDOF\_Group = -1);}
\indent {\em virtual int numberDOF(ID \&lastDOF\_Groups);}
\indent {\em virtual int sendSelf(int commitTag, Channel \&theChannel);} 
\indent {\em virtual int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);}
\indent // Protected Methods
\indent {\em AnalysisModel *getAnalysisModelPtr(void) const;}
\indent {\em GraphNumberer *getGraphNumbererPtr(void) const;}


\noindent {\bf Constructors}
\indent {\em DOF\_Numberer(GraphNumberer \&theGraphNumberer);} 
//! The integer {\em NUMBERER\_TAG\_DOF\_Numberer}
(defined in \f$<\f$classtags.h\f$>\f$) is passed to the MovableObject classes
//! constructor. Sets the GraphNumberer to be used in the numbering to {\em
//! theGraphNumberer()}.

\indent {\em DOF\_Numberer(int classTag);} 
//! Provided for subclasses. The integer \p classTag is passed to the
//! MovableObject classes constructor.


\indent {\em DOF\_Numberer();} 
//! Provided for FEM\_ObjectBroker. The integer {\em NUMBERER\_TAG\_DOF\_Numberer}
(defined in \f$<\f$classtags.h\f$>\f$) is passed to the MovableObject classes
//! constructor. Sets the GraphNumberer to be used in the numbering to be
\f$0\f$, the GraphNumberer object is created and linked in a {\em
//! recvSelf()} method invocation.


\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$DOF\_Numberer();} 
//! Does nothing.

\noindent {\bf Public Methods }
\indent {\em void setLinks(AnalysisModel \&theModel);}
//! Invoked to set a link to the AnalysisModel from which the
//! DOF\_Numberer will number (provide the equation number in the
//! SystemOfEqn object) the degrees-of-freedoms in each DOF\_Group
//! objects.  

\indent {\em virtual int numberDOF(int lastDOF\_Group = -1);}
//! Invoked to assign the equation numbers to the dofs in the DOF\_Groups
//! and the FE\_Elements, ensuring that the dof's in the DOF\_Group whose
//! tag is given by {\em lastDOF\_Group} are numbered last in a \f$-2\f$ or
\f$-3\f$ group. The initial values of these equation numbers have been set
//! by the ConstraintHandler object to be \f$-1\f$, \f$-2\f$ or \f$-3\f$, all dofs
//! with a \f$-3\f$ are to be assigned higher equation numbers than those
//! assigned a \f$-2\f$. To set the \p numEqn in the AnalysisModel and to
//! return the number of equations \p numEqn if successful, a negative
//! number if not.  

//! This base class performs the ordering by getting an ID containing the
//! ordered DOF\_Group tags, obtained by invoking {\em
//! number(theModel-\f$>\f$getDOFGroupGraph(), lastDOF\_Group)} on the
//! GraphNumberer, \p theGraphNumberer, passed in the constructor. The
//! base class then makes two passes through the DOF\_Group objects in the
//! AnalysisModel by looping through this ID; in the first pass assigning the
//! equation numbers incrementally to any degree-of-freedom marked with a
\f$-2\f$ and in the second pass assigning the equation numbers
//! incrementally to any degree-of-freedom marked with a \f$-3\f$. It then
//! iterates through the FE\_Elements in the AnalsisModel invoking {\em
//! setID()} on each object. Finally {\em setNumEqn(numEqn)} is invoked on
//! the AnalysisModel. Return \p numEqn if successful, a warning
//! message and a negative number is returned if an error occurs; \f$-1\f$ is
//! returned if setLinks() has not yet been invoked, \f$-2\f$ if no
//! GraphNumberer was passed in the constructor, \f$-3\f$ if the
//! number of {\em DOF\_Groups} in AnalysisModel and size of ID returned
//! are not the same, and a \f$-4\f$ if there is no DOF\_Group corresponding
//! to one of the tags given in the ID.



\indent {\em virtual int numberDOF(ID \&lastDOF\_Groups);}
//! Invoked to assign the equation numbers to the dofs in the DOF\_Groups
//! and the FE\_Elements, ensuring that the dof's in the DOF\_Groups whose
//! tag is given in {\em lastDOF\_Groups} are numbered last in a \f$-2\f$ or
\f$-3\f$ group. The initial values of these equation numbers have been set
//! by the ConstraintHandler object to be \f$-1\f$, \f$-2\f$ or \f$-3\f$, all dofs
//! with a \f$-3\f$ are to be assigned higher equation numbers than those
//! assigned a \f$-2\f$. To set the \p numEqn in the AnalysisModel and to
//! return the number of equations \p numEqn if successful, a negative
//! number if not.  

//! This method in the base class is almost identical to the one just
//! described. The only difference is that the ID identifying the order of
//! the DOF\_Groups is obtained by invoking {\em
//! number(theModel-\f$>\f$getDOFGroupGraph(), lastDOF\_Groups)} on the
//! GraphNumberer.



\indent {\em virtual int sendSelf(Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);} 
//! The DOF\_Numberer sends the class identifier and database tag of the
//! GraphNumberer in a ID to the channel, if no GraphNumberer is
//! associated a \f$-1\f$ is sent as the class tag. The object then invokes
//! sendSelf() on the GraphNumberer. 

\indent {\em virtual int recvSelf(int commitTag, Channel \&theChannel, FEM\_ObjectBroker \&theBroker);}
//! The DOF\_Numberer receives the class identifier and database tag of
//! the GraphNumberer in an ID from the channel, if no GraphNumberer is
//! associated a \f$-1\f$ is received. The DOF\_Numberer will then ask {\em
//! theBroker} for a GraphNumberer with that class identifier, it sets the
//! database tag for the GraphNumberer and it then invokes {\em
//! recvSelf()} on that GraphNumberer.  

\noindent {\bf Protected Methods}
\indent {\em AnalysisModel *getAnalysisModelPtr(void) const;}
//! A const member function to return the AnalysisModel object associated with
//! the DOF\_Numberer, \p theModel.

\indent {\em GraphNumberer *getGraphNumbererPtr(void) const;}
//! A const member function to return the GraphNumberer object associated with
//! the DOF\_Numberer, \p theGraphNumberer.







