%File: ~/OOP/system_of_eqn/linearSOE/profileSPD/ProfileSPDLinSOE.tex
%What: "@(#) ProfileSPDLinSOE.tex, revA"

\noindent {\bf ProfileSPDLinSOE}

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$ProfileSPDLinSOE.h\f$>\f$

\noindent {\bf Class Decleration}
\indent class ProfileSPDLinSOE: public LinearSOE

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent SystemOfEqn
\indent\indent\indent LinearSOE
\indent\indent\indent\indent {\bf ProfileSPDLinSOE}

\noindent {\bf Description}
\indent ProfileSPDLinSOE is class which is used to store a symmetric
//! system of equations using a profile storage scheme. The upper
//! triangular part of \f$A\f$ is stored in a 1d double array with the diagonals of
\f$A\f$ located at positions given by an integer array \f$iLoc\f$. 
//! For example when \f$n=5\f$ and \f$A\f$ as shown below:

\f$\f$
\left[
\begin{array}{ccccc}
//! a_{0,0} & a_{0,1}  & 0 & 0 & a_{0,4}
//! a_{1,0} & a_{1,1} & a_{1,2} & a_{1,3} & 0
//! a_{2,0} & a_{2,1} & a_{2,2} & a_{2,3} & a_{2,4}
0 & a_{3,1} & a_{3,2} & a_{3,3} & a_{3,4}
0 & 0 & a_{4,2} & a_{4,3} & a_{4,4}
\end{array}
\right] 
\f$\f$

//! this is stored using:


\f$\f$ A =
\left[
\begin{array}{cccccccccccccccccccc}
//! a_{0,0} & a_{0,1}  & a_{1,1} & a_{1,2} & a_{2,2} & a_{1,3} &
//! a_{2,3} & a_{3,3} & a_{0,4} & 0 & a_{2,4} & a_{3,4} & a_{4,4}
\end{array}
\right] 
\f$\f$

//! and 

\f$\f$ iLoc =
\left[
\begin{array}{cccccccccccccccccccc}
1 & 3 & 5 & 8 & 13
\end{array}
\right] 
\f$\f$
//! Note \f$iLoc\f$ stores the diagonal locations using Fortran indexing. This
//! is to facilitate calls to Fortran libraries, e.g. Digital's DXML.
//! The \f$x\f$ and \f$b\f$ vectors are stored in 1d double arrays of length \f$N\f$.


\noindent {\bf Interface}
\indent\indent {// Constructors}
\indent\indent {\em ProfileSPDLinSOE(Solver \&theSolver);}
\indent\indent {\em ProfileSPDLinSOE(int N, int *iLoc,
		  ProfileSPDLinSolver \&theSolver);}
\indent\indent {// Destructor}
\indent\indent {\em  \f$\tilde{}\f$ProfileSPDLinSOE();}
\indent\indent {// Public Methods }
\indent\indent {\em  int setProfileSPDSolver(ProfileSPDLinSolver \&newSolver);}
\indent\indent {\em int setSize(const Graph \&theGraph) =0; }
\indent\indent {\em int getNumEqn(void) =0; }
\indent\indent {\em int addA(const Matrix \&theMatrix, const ID \& loc,
//! doublefact = 1.0) =0;}
\indent\indent {\em int addB(const Vector \& theVector, const ID \& loc,
//! double fact = 1.0) =0;}
\indent\indent {\em int setB(const Vector \& theVector, 
//! double fact = 1.0) =0;}
\indent\indent {\em void zeroA(void) =0;}
\indent\indent {\em void zeroB(void) =0;}
\indent\indent {\em const Vector \&getX(void) = 0;}
\indent\indent {\em const Vector \&getB(void) = 0;}
\indent\indent {\em double normRHS(void) =0;}
\indent\indent {\em void setX(int loc, double value) =0;}
\indent\indent {\em int sendSelf(int commitTag, Channel \&theChannel);} 
\indent\indent {\em int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);} 

\noindent {\bf Constructors}
\indent {\em ProfileSPDLinSOE(Solver \&theSolver);}
//! The \p solver and a unique class tag (defined in \f$<\f$classTags.h\f$>\f$)
//! are passed to the LinearSOE constructor. The system size is set
//! to \f$0\f$ and the matrix \f$A\f$ is marked as not having been factored. Invokes
{\em setLinearSOE(*this)} on the \p solver. No memory is
//! allocated for the 1d arrays.  


{\em ProfileSPDLinSOE(int N, int *newIloc,
		  ProfileSPDLinSolver \&theSolver); }
//! The \p solver and a unique class tag (defined in \f$<\f$classTags.h\f$>\f$)
//! are passed to the LinearSOE constructor. The system size is set
//! to \f$N\f$ and the matrix \f$A\f$ is marked as not having been
//! factored or condensed. Obtains memory from the heap for the 1d arrays storing the
//! data for \f$A\f$, \f$x\f$, \f$b\f$ and \f$iLoc\f$ and stores the size of these arrays. If not
//! enough memory is available for these arrays a warning message is
//! printed and the system size is set to \f$0\f$. The size of \f$A\f$ is given
//! by \f$newIloc(N-1)\f$, if this is not a valid address in \p newIloc a
//! segmentation fault or erronious results will result. The contents of
\f$iLoc\f$ are set equal to those of \p newIloc. Invokes {\em
//! setLinearSOE(*this)} and setSize() on \p solver,
//! printing a warning message if setSize() returns a negative
//! number. Also creates Vector objects for \f$x\f$ and \f$b\f$ using the {\em
(double *,int)} Vector constructor.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$ProfileSPDLinSOE();} 
//! Calls delete on any arrays created.

\noindent {\bf Public Methods }
\indent {\em  int setProfileSPDSolver(ProfileSPDLinSolver \&newSolver);}
//! Invokes {\em setLinearSOE(*this)} on \p newSolver.
//! If the system size is not equal to \f$0\f$, it also invokes setSize()
//! on \p newSolver, printing a warning and returning the returned value if this
//! method returns a number less than \f$0\f$. Finally it returns the result
//! of invoking the LinearSOE classes setSolver() method.

\indent {\em int getNumEqn(void) =0; }
//! A method which returns the current size of the system.

\indent {\em  int setSize(const Graph \&G); } 
//! The size of the system is determined by looking at the adjacency ID of
//! each Vertex in the Graph object \p G. This is done by first
//! determining the column height for each Vertex \f$i\f$ in \p G, done by
//! setting \f$iLoc(i)\f$ equal to \f$0\f$ and then checking for each Vertex
//! in \p G, whether any of the vertex tags in the Vertices adjacency
//! ID results in \f$iLoc(i)\f$ being increased. Knowing the col height of
//! each column, the values of \p iLoc can be determined. Knowing {\em
//! iLoc} and the size of the system (the number of Vertices in \p G, 
//! a check to see if the previously allocated 1d arrays for \f$A\f$, \f$x\f$ and
\f$b\f$ are large enough. If the memory portions allocated for the 1d
//! arrays are not big enough, the old space is returned to the heap and
//! new space is allocated from the heap. Printins a warning message if
//! not enough memory is available on the heap for the 1d arrays and
//! returns a \f$-1\f$. If memory is available, the components of the arrays
//! are zeroed and \f$A\f$ is marked as being unfactored. If the system size
//! has increased, new Vector objects for \f$x\f$ and \f$b\f$ using the {\em
(double *,int)} Vector constructor are created. Finally, the result of 
//! invoking setSize() on the associated Solver object is
//! returned. 


\indent {\em int addA(const Matrix \&M, const ID \& loc,
//! doublefact = 1.0) =0;}
//! First tests that \p loc and \p M are of compatable sizes; if not
//! a warning message is printed and a \f$-1\f$ is returned. The LinearSOE
//! object then assembles \p fact times the Matrix {\em 
//! M} into the matrix \f$A\f$. The Matrix is assembled into \f$A\f$ at the
//! locations given by the ID object \p loc, i.e. \f$a_{loc(i),loc(j)} +=
//! fact * M(i,j)\f$. If the location specified is outside the range,
//! i.e. \f$(-1,-1)\f$ the corrseponding entry in \p M is not added to
\f$A\f$. If \p fact is equal to \f$0.0\f$ or \f$1.0\f$, more efficient steps
//! are performed. Returns \f$0\f$.


{\em int addB(const Vector \& V, const ID \& loc,
//! double fact = 1.0) =0;}
//! First tests that \p loc and \p V are of compatable sizes; if not
//! a warning message is printed and a \f$-1\f$ is returned. The LinearSOE
//! object then assembles \p fact times the Vector \p V into
//! the vector \f$b\f$. The Vector is assembled into \f$b\f$ at the locations
//! given by the ID object \p loc, i.e. \f$b_{loc(i)} += fact * V(i)\f$. If a
//! location specified is outside the range, e.g. \f$-1\f$, the corresponding
//! entry in \p V is not added to \f$b\f$. If \p fact is equal to \f$0.0\f$,
\f$1.0\f$ or \f$-1.0\f$, more efficient steps are performed. Returns \f$0\f$.


{\em int setB(const Vector \& V, double fact = 1.0) =0;}
//! First tests that \p V and the size of the system are of compatable
//! sizes; if not a warning message is printed and a \f$-1\f$ is returned. The
//! LinearSOE object then sets the vector \p b to be \p fact times
//! the Vector \p V. If \p fact is equal to \f$0.0\f$, \f$1.0\f$ or \f$-1.0\f$,
//! more efficient steps are performed. Returns \f$0\f$. 

{\em void zeroA(void) =0;}
//! Zeros the entries in the 1d array for \f$A\f$ and marks the system as not
//! having been factored.

{\em void zeroB(void) =0;}
//! Zeros the entries in the 1d array for \f$b\f$.

{\em const Vector \&getX(void) = 0;}
//! Returns the Vector object created for \f$x\f$.

{\em const Vector \&getB(void) = 0;}
//! Returns the Vector object created for \f$b\f$.

{\em double normRHS(void) =0;}
//! Returns the 2-norm of the vector \f$x\f$.

{\em void setX(int loc, double value) =0;}
//! If \p loc is within the range of \f$x\f$, sets \f$x(loc) = value\f$.

\indent {\em int sendSelf(int commitTag, Channel \&theChannel);} 
//! Returns \f$0\f$. The object does not send any data or connectivity
//! information as this is not needed in the finite element design.

\indent {\em int recvSelf(int commitTag, Channel \&theChannel, FEM\_ObjectBroker
\&theBroker);} 
//! Returns \f$0\f$. The object does not receive any data or connectivity
//! information as this is not needed in the finite element design.




