%File: ~/OOP/system_of_eqn/linearSOE/bandSPD/ProfileSPDLinDirectSkypackSolver.tex
%What: "@(#) ProfileSPDLinDirectSkypackSolver.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/system\_of\_eqn/linearSOE/profileSPD/ProfileSPDLinDirectSkypackSolver.h\f$>\f$

\noindent {\bf Class Decleration}
\indent class ProfileSPDLinDirectSkypackSolver: public LinearSOESolver

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent  Solver
\indent\indent\indent LinearSOESolver
\indent\indent\indent\indent ProfileSPDLinSolver
\indent\indent\indent\indent\indent {\bf ProfileSPDLinDirectSkypackSolver}

\noindent {\bf Description}
\indent A ProfileSPDLinDirectSkypackSolver object can be constructed
//! to solve a ProfileSPDLinSOE object. It does this by direct means using
//! the routines supplied in the SKYPACK library, a library which uses the
//! BLAS levels 1,2 and 3 for the factorization and substitution. 

//! The routines in SKYPACK require a number of work areas: {\em int
//! block[3]} and {\em double invD[size]}. In addition, to allow the use
//! of the BLAS 2 and 3, work areas {\em double rw[mRows x mCols]}, {\em
//! double tw[mRows x mRows]} and {\em int index[max(mCols,mRows)]} are 
//! created.


\noindent {\bf Interface}
\indent\indent Constructors
\indent\indent {\em ProfileSPDLinDirectSkypackSolver();} 
\indent\indent {\em ProfileSPDLinDirectSkypackSolver(int mCols, mRows);}
\indent\indent Destructor
\indent\indent {\em  \f$\tilde{}\f$ProfileSPDLinDirectSkypackSolver();}
\indent\indent Public Methods
\indent\indent {\em int solve(void);}
\indent\indent {\em  int setSize(void);}
\indent\indent {\em int sendSelf(int commitTag, Channel \&theChannel);} 
\indent\indent {\em int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker);} 


\noindent {\bf Constructors}
\indent {\em ProfileSPDLinDirectSkypackSolver();}
//! A unique class tag (defined in \f$<\f$classTags.h\f$>\f$) is passed to the
//! ProfileSPDLinSolver constructor. Sets \p mCols and \p mRows
//! equal to \f$0\f$ and does not try and allocate any memory for the work arrays.

\indent {\em ProfileSPDLinDirectSkypackSolver(int mCols, int mRows);}
//! A unique class tag (defined in \f$<\f$classTags.h\f$>\f$) is passed to the
//! ProfileSPDLinSolver constructor. Sets \p mCols and \p mRows and
//! allocates space in memory for the work arrays \p rw, \p tw and
\p index. If not enough memory is available in memory, \p mCols
//! and \p mRows is set equal to \f$0\f$ and an error message is printed.

\noindent {\bf Destructor}
\indent {\em \f$\tilde{}\f$ProfileSPDLinDirectSkypackSolver();} 
//! Invokes delete on any work areas that have been constructed.

\noindent {\bf Public Member Functions }
\indent {\em  int solve(void);}
//! The solver first copies the B vector into X and then solves the
//! BandSPDLinSOE system. If the matrix has not been factored, the matrix
//! is first factored using the SKYPACK routine skysf2(), if {\em
//! mCols} and \p mRows equal \f$0\f$, or skypf2()}. {\em skysf2() is
//! a routine which uses the BLAS level 1 routines, skypf2() is a
//! routine which uses BLAS levels 2 and 3. If skypf2()
//! has been called, \p invD is set up. Once the matrix has been
//! factored, skyss() is called. If the solution is sucessfully
//! obtained, i.e. the skyss() routine returns \f$0\f$ in the INFO
//! argument, \f$0\f$ is returned, otherwise it prints a warning message and
//! returns INFO. The solve process changes \f$A\f$ and \f$X\f$.   

\indent {\em  int setSize(void);}
//! Is responsible for setting the \p block information required by the
//! SKYPACK routines (block[0]=1; block[1]=size, block[2]=1) and for
//! creating space for the \p invD work array. Returns \f$0\f$ if
//! successfull, otherwise a warning message is printed and a \f$-1\f$ is
//! returned.

\indent {\em  int sendSelf(int commitTag, Channel \&theChannel);} 
//! Does nothing but return \f$0\f$.

\indent {\em  int recvSelf(int commitTag, Channel \&theChannel, FEM\_ObjectBroker
\&theBroker);} 
//! Does nothing but return \f$0\f$.







