% File: ~/OOP/analysis/model/DOF_Group.tex 
%What: "@(#) DOF_Group.tex, revA"

//! NEED A GETTANGFORCE() LIKE FE\_ELEMENT FOR ELE\_BY\_ELE SOLVERS. KEEP A
//! POINTER TO LAST INTEGRATOR.

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/analysis/dof\_grp/DOF\_Group.h\f$>\f$

\noindent {\bf Class Decleration}
\indent class DOF\_Group;

\noindent {\bf Description}
\indent DOF\_Group is a base class. An object of type DOF\_Group
//! represents an unconstrained node of the domain in the model. Each node
//! in the domain is associated with one DOF\_Group. DOF\_Groups are
//! called upon in the analysis to provide their contributions of
//! unbalanced load to the system of equations. Subclasses are used by the
//! constraint handler to to introduce new dofs into the analysis. 

//! The DOF\_Group is responsible for providing operations to set and
//! access the mapping betwwen equation numbers and DOFs, allowing the
//! Integrator to from the tangent (if nodal masses in transient problem)
//! and unbalanced load information, and for setting and obtaining the
//! nodal trial response quantities.


\noindent {\bf Class Interface}
\indent // Constructors
\indent {\em DOF\_Group(int tag, Node *theNode);}
\indent {\em DOF\_Group(int tag, int numDOF);}
\indent // Destructor
\indent {\em virtual~ \f$\tilde{}\f$DOF\_Group();}
\indent // Public Methods - Mapping
\indent {\em virtual void setID(int dof, int value);}
\indent {\em virtual void setID(const ID \&values);}
\indent {\em virtual const ID \&getID(void) const;}
\indent {\em virtual int getTag(void) const;}
\indent {\em virtual int getNumDOF(void) const;}
\indent {\em virtual int getNumFreeDOF(void) const;}
\indent {\em virtual int getNumConstrainedDOF(void) const;}
\indent // Public Methods - Tangent \& Residual 
\indent {\em virtual const Matrix \&getTangent(Integrator *theIntegrator)}
\indent {\em virtual void zeroTangent(void);}
\indent {\em virtual void addMtoTang(double factt);}
\indent {\em virtual const Vector \&getUnbalance(Integrator
*theIntegrator); } 
\indent {\em virtual void zeroUnbalance(void);}
\indent {\em virtual void  addPtoUnbalance(double fact = 1.0);}
\indent {\em virtual void addMtoTang(const Vector \&\f$\ddot u\f$,
//! double fact);} 
\indent // Public Methods - Node Response
\indent {\em virtual const Vector \&getCommittedDisp(void);}
\indent {\em virtual const Vector \&getCommittedVel(void);}
\indent {\em virtual const Vector \&getCommittedAccel(void);}
\indent {\em virtual int setNodeDisp(const Vector \&u);}
\indent {\em virtual int setNodeVel(const Vector \&\f$\dot u\f$);}
\indent {\em virtual int setNodeAccel(const Vector \&\f$\ddot u\f$);}
\indent {\em virtual int incrNodeDisp(const Vector \&u);}
\indent {\em virtual int incrNodeVel(const Vector \&\f$\dot u\f$);}
\indent {\em virtual int incrNodeAccel(const Vector \&\f$\ddot u\f$);}



\noindent {\bf Constructors}
\indent {\em DOF\_Group(int tag, Node *theNode);}
//! Constructs a  DOF\_Group with an associated node given by {\em
//! theNode} and a tag given by \p tag. During construction it
//! determines the number of unknown dofs from the node, \p numDOF, and
//! creates an Vector object to hold the unbalance information and an ID
//! object to hold mapping between degrees-of-freedom and equation
//! numbers. All values of the ID are set to \f$-2\f$ initially. It creates a
//! Vector to hold the unbalance loads and an ID to 
//! hold the mapping information, both of size equal to the number of
//! unknowns. If the size of the Vector or the ID is \f$0\f$, i.e. ran out of
//! memory, a warning message is orinted and \p numDOF set to \f$0\f$.

{\em DOF\_Group(int tag, int numDOF);}
//! Provided for subclasses. Constructs a  DOF\_Group with the number of
//! unknown dofs given by \p numDOF and a tag given by \p tag. No
//! Node is associated with this DOF\_Group. Creates a Vector of size {\em
//! numDOF} to hold the unbalance information and an ID object to
//! hold mapping between degrees-of-freedom and equation numbers. All
//! values of the ID are set to \f$-2\f$ initially. If
//! the size of the Vector or the ID is \f$0\f$, i.e. ran out of memory, a
//! warning message is orinted and \p numDOF set to \f$0\f$.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$DOF\_Group();}
//! Invokes the destructor on the Vector created to hold the unbalance.

\noindent {\bf Public Methods - Equation Numbers}
\indent {\em virtual void setID(int dof, int value);}
//! Operation to set the equation number of the {\em dof'th} DOF in the
//! DOF\_Group to \p value. In this base class, if \p index is a
//! valid location, \f$0 < index < numDOF-1\f$ the operator {\em (index) =
//! value} is invoked on the ID object created for the mapping, otherwise
//! an error message is printed. 


{\em virtual void setID(const ID \&values);}
//! An operation to set the equation numbers of all the DOFs in the
//! DOF\_Group to be those given in the ID \p values. In this base
//! class, the operator {\em = values} is invoked on the ID object created
//! for the mapping. 

{\em virtual const ID \&getID(void) const;}
//! A const member function which returns in an ID object the equation
//! numbers associated with the degrees-of-freedom in the DOF\_Group. The
//! size of the ID object is equal to the number of degrees-of-freedom
//! represented by the DOF\_Group. This base class returns the ID created
//! for the mapping. 

{\em virtual int getTag(void) const;}
//! A const member function which returns the integer identifier {\em
//! tag} passed in the constructor.

{\em virtual int getNumDOF(void) const;}
//! Returns the total number of DOFs in the DOF\_Group. This base class
//! returns \p numDOF.

{\em virtual int getNumFreeDOF(void) const;}
//! Returns the total number of \p free DOFs in the DOF\_Group,
//! i.e. number of dof in the group who have not been assigned a negative
//! equation number. Determines this by looping through the ID containing
//! the mapping information.

{\em virtual int getNumConstrainedDOF(void) const;}
//! Returns the total number of \p constrained DOFs in the DOF\_Group,
//! i.e. number of dof in the group who have been assigned a negative
//! equation number. Determines this by looping through the ID containing
//! the mapping information.


\noindent {\bf Public Methods - Nodal Tangent contributions
//! for transient} 
\indent {\em virtual const Matrix \&getTangent(void)}
//! Returns the current tangent matrix for the DOF\_Group. If no tangent
//! matrix has been created, one of size(numDOF,numDOF) is constructed.
//! If not enough space is available for a new tangent matrix, an error
//! message is printed and an error Matrix is returned..  

\indent {\em virtual void formTangent(Integrator *theIntegrator)}
//! To form the DOF\_Groups tangent. Invokes {\em formNodTangent(this)} on
//! The Integrator object \p theIntegrator.

\indent {\em virtual void zeroTangent(void);}
//! To zero the tangent. If a tangent matrix has been allocated,
//! will invoke Zero() on the tangent matrix.

\indent {\em virtual void addMtoTang(double fcat);}
//! To add \p fact times the nodal mass matrix to the tangent
//! matrix. In this base class, if a Matrix to store the tangent has not
//! yet been created, one of size (numDOF,numDOF) is now created; if
//! construction of this matrix fails an error is printed and an error
//! Matrix is returned. Invokes {\em addMatrix(theNode-\f$>\f$getMass,fact)}
//! on the Matrix, printing a warining message if this method returns a
\f$negative\f$ value. If no Node is associated with the DOF\_Group an
//! error message is printed and nothing is done.

\noindent {\bf Public Methods - Node Unbalance}
\indent {\em virtual void zeroUnbalance(void);}
//! To zero the unbalance vector. Invokes Zero() on the vector
//! object used to store the unbalance information.

{\em virtual void formUnbalance(Integrator *theIntegrator); }
//! Causes the DOF\_Group to form its contribution to the residual. Invokes
{\em formNodUnbalance(this)} on \p theIntegrator.

{\em virtual const Vector \&getUnbalance() const; }
//! Returns the vector holding the unbalance.

{\em virtual void  addPtoUnbalance(double fact = 1.0);}
//! Adds the product of the unbalanced load at the node and \p fact to
//! the unbalance vector. A warning message is printed and nothing is done
//! if no node is associated with the DOF\_Group, otherwise {\em
//! addVector(theNode-\f$>\f$getUnbalancedLoad(),fact)} is invoked on the
//! unbalance vector and a warning message is printed if this method
//! returns a negative value. 

{\em virtual void addMtoTang(const Vector \&\f$\ddot u\f$,
//! double fact);} 
//! Adds \p fact times the product of the associated nodes mass matrix
//! and the Vector \f$\ddot u\f$ to the unbalance vector. A warning message is
//! printed and nothing is done if no node is associated with the
//! DOF\_Group.

\noindent {\bf Public Methods - Node Response}
\indent {\em virtual const Vector \&getCommittedDisp(void);}
//! To return the committed displacement at the node. Returns the result
//! of invoking getDisp() on the Node. If there is no associated
//! node object, an error message is printed and an error Vector is
//! returned.


\indent {\em virtual const Vector \&getCommittedVel(void);}
//! To return the committed velocity at the node. Returns the result of
//! invoking getVel() on the Node. If there is no associated node
//! object, an error message is printed and an error Vector is returned.

\indent {\em virtual const Vector \&getCommittedAccel(void);}
//! To return the committed velocity at the node. Returns result of
//! invoking  getAccel() on the Node. If there is no associated node
//! object, an error message is printed and an error Vector is returned.

\indent {\em virtual int setNodeDisp(const Vector \&u);}
//! This operation sets the value of the nodal trial displacement at the
//! associated node. The vector \p u is of size equal to the number of
//! equations in the AnalysisModel (this is not checked by the DOF\_Group).
//! The DOF\_Group object accesses the contents of the Vector \p u using
//! its assigned equation numbers. If a \f$-1\f$ exists for a DOF a \f$0\f$ value
//! is set as the corresponding nodal displacement for the node. Creates a Vector
//! to store the correct components of \p u, then invokes setTrialDisp()
//! on the node object; if no node object an error message is
//! printed. CURRENTLY THERE IS NO CHECK TO SEE IF myID(i) DOES NOT OVERFLOW
//! ADMISSABLE VALUES IN U - THIS NEEDS TO CHANGE


{\em virtual int setNodeVel(const Vector \&\f$u^{.}\f$);}
//! This operation sets the value of the nodal trial velocity at the
//! associated node. The vector {\em \f$u^{.}\f$} is of size equal to the number of
//! equations in the AnalysisModel (this is not checked by the DOF\_Group).
//! The DOF\_Group object accesses the contents of the Vector {\em \f$u^{.}\f$} using
//! its assigned equation numbers. If a \f$-1\f$ exists for a DOF a \f$0\f$ value
//! is set as the corresponding nodal velocity for the node.  Creates a Vector
//! to store the correct components of \p u, then invokes setTrialVel()
//! on the node object; if no node object an error message is printed. 
//! MUST CHANGE AS setNodeDisp

{\em virtual int setNodeAccel(const Vector \&\f$u^{..}\f$);}
//! This operation sets the value of the nodal trial acceleration at the
//! associated node. The vector {\em \f$u^{..}\f$} is of size equal to the number of
//! equations in the AnalysisModel (this is not checked by the DOF\_Group).
//! The DOF\_Group object accesses the contents of the Vector {\em \f$u^{..}\f$} using
//! its assigned equation numbers. If a \f$-1\f$ exists for a DOF a \f$0\f$ value
//! is set as the corresponding nodal acceleration for the node. The return
//! value is as outlined above for {\em setNodeAccel().} Creates a Vector
//! to store the correct components of \p u, then invokes setTrialAccel()
//! on the node object; if no node object an error message is printed. 
//! MUST CHANGE AS setNodeDisp



\indent {\em virtual int incrNodeDisp(const Vector \&u);}
\indent {\em virtual int incrNodeVel(const Vector \&\f$u^{.}\f$);}
\indent {\em virtual int incrNodeAccel(const Vector \&\f$u^{..}\f$);}
//! These methods are similar to those three just outlined, the only
//! difference being that the trial response quantities at the nodes are
//! incremented, not set, with the values now given; this is done by
//! invoking incrTrialDisp()}, {\em incrTrialVel(), and {\em 
//! incrTrialAccel()} respectivly on the nodes. MUST ALL CHANGE AS
//! setNodeDisp 






