% File: ~/utility/tagged/TaggedObject.tex 
% What: "@(#) TaggedObject.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/utility/tagged/TaggedObject.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class TaggedObject

\noindent {\bf Class Hierarchy}
\indent  {\bf TaggedObject}
\indent\indent  {DomainComponent}
\indent\indent  {Vertex}

\noindent {\bf Description}
\indent TaggedObject is used as a base class to represent all classes
//! that may have a integer identifier, a tag, to identify the object. It
//! is used in the framework as a base class for many classes, for example
//! DomainComponent and Vertex. The class is provided so that container
//! classes can be written to store objects and provide access to
//! them. This saves us rewriting container classes for each type of 
//! object. (templates will be able to provide this functionality when
//! they are provided with all compilers). 

\noindent {\bf Class Interface}
\indent // Constructor
\indent {\em TaggedObject(int tag);}
\indent // Destructor
\indent {\em virtual~ \f$\tilde{}\f$TaggedObject();}
\indent // Public Methods
\indent {\em int getTag(void) const;} 
\indent {\em virtual void Print(OPS_Stream \&s) =0;}
\indent {\em friend OPS_Stream \&operator\f$<<\f$(OPS_Stream \&s,
//! TaggedObject \&m);} 
\indent // Protected Methods
\indent {\em virtual void setTag(int newTag);}

\noindent {\bf Constructors}
\indent {\em TaggedObject(int tag);}
//! Constructs a TaggedObject with a tag given by \p tag. The tag of
//! a component is some unique means of identifying the component among
//! like components, i.e. the tag of a node would be its unique node number.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$TaggedObject();}
//! Does nothing. Provided so the concrete subclasses destructors will be
//! called.

\noindent {\bf Public Member Functions}
\indent {\em int getTag(void) const;}
//! Returns the tag associated with the object. This function is inlined
//! for performance.

\indent {\em virtual void Print(OPS_Stream \&s, int flag =0) =0;}
//! A pure virtual function. The component is to output itself to the output
//! stream \p s. The integer \p flag can be used to select just what
//! should be output, by default \f$0\f$ is passed.

\indent {\em friend OPS_Stream \&operator\f$<<\f$(OPS_Stream \&s, TaggedObject \&m);}  
//! Invokes {\em Print(s)} on the TaggedObject \p m.

\noindent {\bf Protected Member Functions}
\indent {\em virtual void setTag(int newTag);}
//! Sets the tag of the object to be \p newTag. It is provided so that
//! MovableObjects can set their tag in recvSelf().







