% File: ~/domain/domain/loadBalancer/SwapHeavierToLighterNeighbours.tex 
% What: "@(#) SwapHeavierToLighterNeighbours.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/domain/loadBalancer/SwapHeavierToLighterNeighbours.h\f$>\f$

\noindent {\bf Class Decleration}
\indent class SwapHeavierToLighterNeighbours: public LoadBalancer

\noindent {\bf Class Hierarchy}
\indent  LoadBalancer
\indent\indent {\bf SwapHeavierToLighterNeighbours}


\noindent {\bf Description}
\indent A SwapHeavierToLighterNeighbours is an object used to balance a
//! PartitionedDomain. It does this by shedding the boundary vertices on
//! the heaviest loaded partition (subdomain).

\noindent {\bf Class Interface}
\indent\indent  // Constructors
\indent\indent {\em SwapHeavierToLighterNeighbours(); } 
\indent\indent {\em SwapHeavierToLighterNeighbours(double
//! factorGreater, int numReleases);}
\indent\indent // Destructor
\indent\indent {\em virtual~ \f$\tilde{}\f$SwapHeavierToLighterNeighbours();}
\indent\indent // Public Methods
\indent\indent {\em virtual int balance(Graph \&theWeightedGraph) =0;}


\noindent {\bf  Constructors  }
\indent {\em SwapHeavierToLighterNeighbours(); } 
//! Sets \p numRealeases to \f$1\f$ and \p factorGreater to \f$1.0\f$. These
//! are the paramemeters used in the balance() method.

\indent {\em SwapHeavierToLighterNeighbours(double factorGreater, int
//! numReleases);} 
//! Sets the parameters used in the balance() method.

\noindent {\bf Destructor }
\indent {\em virtual~ \f$\tilde{}\f$SwapHeavierToLighterNeighbours();} 
//! Does nothing.

\noindent {\bf  Public Methods}
\indent {\em virtual int balance(Graph \&theWeightedGraph) =0;} 
//! For \p numRelease times the Vertices of \p theWeightedGraph are
//! iterated through. For each Vertex, \f$i\f$, the weight of the Vertex is compared
//! to those of it's adjacent Vertices, \f$other\f$, (this is done by looping through
//! the adjacency ID of the Vertex), if the vertex weight is {\em
//! factorGreater} times greater than the other Vertices load then {\em
//! swapBoundary(i, other)} is invoked on the DomainPartitioner. Returns
\f$0\f$ if successfull, otherwise a negative number and a warning message
//! are returned if either no link has been set to the DomainPartitioner
//! or swapBoundary() returns a negative number.



