%File: ~/OOP/system_of_eqn/linearSOE/LinearSOESolver.tex
%What: "@(#) LinearSOESolver.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/system\_of\_eqn/linearSOE/LinearSOESolver.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class LinearSOESolver: public Solver

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent  Solver
\indent\indent\indent {\bf LinearSOESolver}

\noindent {\bf Description}
\indent LinearSOESolver is an abstract class. A LinearSOESolver object is
//! responsible for solving the LinearSOE object that it is associated
//! with. That is, to find \f$x\f$ such that the matrix equation \f$Ax=b\f$ is
//! satisfied.

\noindent {\bf Interface}
\indent\indent {// Constructor}
\indent\indent {\em LinearSOESolver(int classTag);}
\indent\indent {// Destructor}
\indent\indent {\em virtual~ \f$\tilde{}\f$LinearSOESolver();}
\indent\indent {// Public Methods }
\indent\indent {\em virtual int solve(void) =0;}
\indent\indent {\em virtual int setSize(void) =0;}

\noindent {\bf Constructor}
\indent {\em LinearSOESolver(int classTag);}
//! The integer \p classTag is passed to the Solver.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$LinearSOESolver();} 
//! Does nothing. Provided so the subclasses destructor will be called.

\noindent {\bf Public Methods }
\indent {\em virtual int solve(void) =0;}
//! Causes the LinearSOESolver to solve the system of equations \f$Ax=b\f$ for \f$x\f$.
//! Returns \f$0\f$ if successful , negative number if not; the actual value depending on
//! the type of LinearSOESolver. The result of the solve are to be stored
//! in the \f$x\f$ vector of the LinearSOE by the object.

\indent {\em virtual int setSize(void) =0;}
//! This is invoked by the \p LinearSOE object when setSize() has
//! been invoked on it. Solvers may sometimes need to store additional
//! data that needs to be updated if the size of the system of equation
//! changes.




