%File: ~/OOP/analysis/handler/PlainHandler.tex
%What: "@(#) PlainHandler.tex, revA"

\noindent {\bf Files}
\indent \#include \f$<\tilde{ }\f$/analysis/handler/PlainHandler.h\f$>\f$

\noindent {\bf Class Declaration}
\indent class PlainHandler: public ConstraintHandler

\noindent {\bf Class Hierarchy}
\indent MovableObject
\indent\indent ConstraintHandler
\indent\indent\indent {\bf PlainHandler}

\noindent {\bf Description} 
\indent The PlainHandler class is a class which only deals with
//! homogeneous single point constraints. To do this it creates regular
//! FE\_Element and DOF\_Group objects and enforces the constraints by
//! specifying that degrees-of-freedom which are constrained are not
//! assigned an equation number. Pointers to the DOF\_Group and
//! FE\_Element objects are kept in two arrays.

\noindent {\bf Class Interface}
\indent // Constructor
\indent {\em PlainHandler();}
\indent // Destructor
\indent {\em \f$\tilde{ }\f$PlainHandler();}
\indent // Public Methods
\indent {\em int handle(const ID *nodeToBeNumberedLast
=0);} 
\indent {\em void clearAll(void);}
\indent {\em int sendSelf(int commitTag, Channel \&theChannel); }
\indent {\em int recvSelf(int commitTag, Channel \&theChannel,
//! FEM\_ObjectBroker \&theBroker); }

\noindent {\bf Constructor}
\indent {\em PlainHandler();} 
//! The integer {\em HANDLER\_TAG\_PlainHandler} (defined in
\f$<\f$classTags.h\f$>\f$) is passed to the PlainHandler constructor.

\noindent {\bf Destructor}
\indent {\em virtual~ \f$\tilde{}\f$PlainHandler();} 
//! Invokes the destructor on all the FE\_Element and DOF\_Group objects
//! created in \p handle. Then invokes the destructor on the two
//! pointer arrays.

\noindent {\bf Public Methods }
\indent {\em int handle(const ID *nodeToBeNumberedLast =0);}
//! Determines the number of FE\_Elements and DOF\_Groups needed from the
//! Domain (a one to one mapping between Elements and FE\_Elements and
//! Nodes and DOF\_Groups) Creates two arrays of pointers to store the
//! FE\_elements and DOF\_Groups, returning a warning message and a \f$-2\f$
//! or \f$-3\f$ if not enough memory is available for these arrays. Then the
//! object will iterate through the Nodes of the Domain, creating a
//! DOF\_Group for each node and setting the initial id for each dof to
\f$-2\f$ if no SFreedom\_Constraint exists for the dof, or \f$-1\f$ if a
//! SFreedom\_Constraint exists or \f$-3\f$ if the node identifier is in {\em
//! nodesToBeNumberedLast}. The object then iterates through the Elements
//! of the Domain creating a FE\_Element for each Element, if the Element
//! is a Subdomain setFE\_ElementPtr() is invoked on the Subdomain
//! with the new FE\_Element as the argument. If not enough memory is
//! available for any DOF\_Group or FE\_element a warning message is
//! printed and a \f$-4\f$ or \f$-5\f$ is returned. If any MP\_Constraint objects
//! exist in the Domain a warning message is printed and \f$-6\f$ is
//! returned. If all is successful, the method returns the number of
//! degrees-of-freedom associated with the DOF\_Groups in {\em
//! nodesToBeNumberedLast}. 

{\em void clearAll(void) =0;}
//! Currently this invokes delete on all the FE\_element and DOF\_Group
//! objects created in handle() and the arrays used to store
//! pointers to these objects. FOR ANALYSIS INVOLVING DYNAMIC LOAD
//! BALANCING, RE-MESHING AND CONTACT THIS MUST CHANGE.

{\em int sendSelf(int commitTag, Channel \&theChannel); }
//! Returns \f$0\f$.

{\em int recvSelf(int commitTag, Channel \&theChannel, FEM\_ObjectBroker
\&theBroker); }
//! Returns \f$0\f$. 
